Assessing primitives performance\\ on multi-stage execution

### 2018-05-08 paper.tex
implem comparison + asm comparison figure

\paragraph{Original specification.}  The original primitive took three operands. The two first operands were the two ByteStrings to compare and the third operand is the order table, a ByteArray encoding the order of characters. The default order is the ASCII order, hence each entry corresponds to its index in the byte array (The ByteArray has a size of 256, and accessing the element i, with 0 <= i <= 255, answered i). However, in some case, the program may do case insensitive ByteStrings comparison. In this case, the order ByteArray is different, since some entries answer the same value (for example, the entry for the 41 and 61, respectively characters A and a, both answer the same value). To compare the two ByteStrings, the primitive conceptually iterates byte by byte over the two objects until it reaches the end of one them, comparing for each byte that the value present in the order ByteArray at the index is greater, equal or smaller than the value computed from the other ByteString. If a difference is met, the value returns based on the result, if not, at the end of the iteration the primitive compares the ByteString sizes and returns. Based on the three parameters, the primitive answered 3,2 or 1 depending if the first ByteString was greater, equal or smaller than the second one.

@@

 \cba{Maybe wrte version A, B, C , D , E or a word to represent what they are (Smalltalk) (SmartSyntax) (Plugin) (Slang) (Slang+JIT)
 so you can refer to that to talk about any of these later and not wirte each time : the implem with ths wirtten that wya and exec that way.
 Myabe those could be paragraph titles instead of the following}

@@

\begin{table}
\centering
\begin{tabular}{|c||c|c|c|c|}
  \hline
    & \rot{Engineering time} & \rot{Execution time} & \rot{Readability}  \\
  \hline
  Baseline & + & +++ & +++ \\
  SmartSyntax & ++ & ++ & ++ \\
  SlangPlugin & ++ & ++ & ++ \\
  PureSlang & +++ & + & ++ \\
  Slang+JIT & +++ & + & + \\
  \hline
\end{tabular}
\caption{Trade-offs between implementations: synthetized}
\label{tab:comparisontable}
\end{table}

Each of these implementations has its pros and cons considering the context. In the end, there is always a trade-off between engineering time and performance. In

@@

\cba{Personnellement je ferais une figure, ou tu places tes graphes dedans, et pares tu referes a la figure dans le texte.}

@@

%\cba{Je pense string de taille 0-1 c'est cool pour montrer que le code C va moins vite que pur smalltalk dans ce cas (si c'est vraiment le cas)}
We performed the benchmarks upon 4 different string sizes: 0, 5, 100 and 1000 characters, using 2 different compilers: GCC on one hand and LLVM on the other hand.

The results are shown in terms of relative speed-up towards the baseline implementation, i.e. the pure smalltalk version. 
We can notice some patterns over the results in figures \ref{fig:gcc} and \ref{fig:llvm}:
\begin{itemize}
\item The execution time of the SmartSyntax and Plugin implementations is either very close or greater than the baseline's for the small strings. This can be explained by the high cost of the switching between the C and Smalltalk stacks, as mentionned in Section \ref{sec:VMexec}.
\item The Slang+JIT implementation is more efficient in every case. Nevertheless, 
the performance gap between this implementation and the PureSlang one tends to decrease as the string size increase. The comparison loop \sk{ref to appendix}
\end{itemize} proves to be more and more time-consuming because of the higher number of iterations.

@@

@@

\cba{Show ASM loops and discuss the differences
add appendix with full code}

>>>>

\paragraph{Original specification.}  The original primitive took three operands. The two first operands were the two ByteStrings to compare and the third operand is the order table, a ByteArray encoding the order of characters. The default order is the ASCII order, hence each entry corresponds to its index in the byte array (The ByteArray has a size of 256, and accessing the element i, with 0 $\leq$ i $\leq$ 255, answered i). However, in some case, the program may do case insensitive ByteStrings comparison. In this case, the order ByteArray is different, since some entries answer the same value (for example, the entry for the 41 and 61, respectively characters A and a, both answer the same value). To compare the two ByteStrings, the primitive conceptually iterates byte by byte over the two objects until it reaches the end of one them, comparing for each byte that the value present in the order ByteArray at the index is greater, equal or smaller than the value computed from the other ByteString. If a difference is met, the value returns based on the result, if not, at the end of the iteration the primitive compares the ByteString sizes and returns. Based on the three parameters, the primitive answered 3,2 or 1 depending if the first ByteString was greater, equal or smaller than the second one.

@@

@@

\begin{table}
\centering
\begin{tabular}{|c|c|c||c||c|c|c||c|}
  \hline
    & \rot{readability} & \rot{maintenance} & \rot{Engineering time} & \rot{re-compilation} & \rot{plugin} & \rot{stack switching} & \rot{Execution time} \\
  \hline
  Baseline    & X &   & +   &    &   &   & +++ \\
  SmartSyntax & X &   & ++  &    & X & X & ++\\
  SlangPlugin &   &   & ++  &    & X & X & ++ \\
  PureSlang   &   & X & +++ &  X &   & X & ++\\
  Slang+JIT   &   & X & +++ &  X &   &   & + \\
  \hline
\end{tabular}
\caption{Trade-offs between implementations: summarized}
\label{tab:cmp_implem}
\end{table}

One or the other implementation could be chosen over another one, depending on what one's aiming at: the main trade-off will be between the engineering time and the performance. The table \ref{cmp_implem} summarizes the main pros and cons related to each implementation

@@

@@

%\cba{Je pense string de taille 0-1 c'est cool pour montrer que le code C va moins vite que pur smalltalk dans ce cas (si c'est vraiment le cas)}
We performed the benchmarks upon 4 different string sizes: 0, 5, 100 and 1000 characters, using 2 different compilers: GCC on one hand and LLVM on the other hand. Figure \ref{fig:gcc} and \ref{fig:llvm} show the results in terms of relative speed-up towards the baseline implementation\footnote{i.e. the pure Smalltalk version}, respectively for GCC and LLVM.

We can draw the following patterns from these results:
\begin{itemize}
\item The execution time of the SmartSyntax and Plugin implementations is either very close or greater than the baseline's for the small strings. This can be explained by the high cost of the switching between the C and Smalltalk stacks, as mentionned in Section \ref{sec:VMexec}.
\item The Slang+JIT implementation is more efficient in every case. Nevertheless, 
the performance gap between this implementation and the PureSlang one tends to decrease as the string size increase. The comparison loop \sk{ref to appendix}
\end{itemize} proves to be more and more time-consuming because of the higher number of iterations.

@@

\sk{parler des autres optis qui n'ont pas d'impact négatif sur le temps d'exec ? ou le conserver pour les future work pour des potentielles améliorations?}

@@

\cba{appendix with full code}
